### 2015-12-04 results/tables/build_bacterial_category_phylum_table.Rmd

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=0.1in
classoption: landscape
---

**Supplementary Table 2. Frequency that each bacterial phylum was found across each of the environmental categories.**

```{r results='asis', echo=FALSE, warning=FALSE, eval=TRUE}
require(xtable, quietly=TRUE)
options(xtable.sanitize.text.function=identity)

categories <- c(AE="Aerosol", AQB="Brackish", AQBS="Brackish sediment", AQF="Freshwater", AQFS="Freshwater sediment", AQM="Marine", AQMS="Marine sediment", AQH="Hydrothermal vent", AQI="Ice", AQO="Aquatic other", BD="Digesters", BF="Food-associated", BI="Industrial/mining", BP="Pollution associated", BO="Built other", PR="Plant root", PS="Plant surface", PO="Plant other", SA="Agricultural soil", SD="Desert soil", SP="Permafrost", SO="Other soils", ZV="Vertebrate", ZA="Arthropod", ZN="Other invertebrate", ZO="Other zoological", OT="Other", total="Total")

data <- read.table(file="data/process/phylum_category_counts.tsv", header=T)

bacteria <- data[data$domain=="bacteria",-1]
bacteria <- bacteria[order(bacteria$total, decreasing=T), names(categories)]
colnames(bacteria) <- gsub("total", "Total", colnames(bacteria))
rownames(bacteria) <- gsub("Candidate_division_", "", rownames(bacteria))
rownames(bacteria) <- gsub("_", " ", rownames(bacteria))

total <- apply(bacteria, 2, sum)
bacteria <- rbind(bacteria, Total=total)
bacteria <- cbind(rownames(bacteria), bacteria)

colnames(bacteria) <- c("", categories)
composite <- xtable(bacteria)

align(composite) <- paste0('ll', paste(rep('c', ncol(bacteria)-1), collapse=""))

print(composite, include.rownames=FALSE, include.colnames=TRUE, hline.after = c(-1,0,nrow(composite)-1, nrow(composite)), comment=FALSE, size="tiny", rotate.colnames=TRUE)

```