### 2015-12-04 results/tables/build_environmental_categories_table.Rmd
Create the environmental category description table

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=1.0in
---

**Supplemental Table 1. Description of environmental categories and the criteria used to assign sequences to each category.**

```{r results='asis', echo=FALSE, warning=FALSE, eval=TRUE}
require(xtable, quietly=TRUE)
options(xtable.sanitize.text.function=identity)

categories <- c(AE="Aerosol", AQB="Brackish", AQBS="Brackish sediment", AQF="Freshwater", AQFS="Freshwater sediment", AQM="Marine", AQMS="Marine sediment", AQH="Hydrothermal vent", AQI="Ice", AQO="Aquatic other", BD="Digesters", BF="Food-associated", BI="Industrial/mining", BP="Pollution associated", BO="Built other", PR="Plant root", PS="Plant surface", PO="Plant other", SA="Agricultural soil", SD="Desert soil", SP="Permafrost", SO="Other soils", ZV="Vertebrate", ZA="Arthropod", ZN="Other invertebrate", ZO="Other zoological", OT="Other")

text <- read.table("doc/notebook/ClassificationCode.md", sep=':', row.names=1, stringsAsFactors=FALSE)

text <- text[names(categories),]

text$V2 <- categories

text$V3 <- gsub("^ +", "", text$V3)
text$V3 <- gsub("^(.)", "\\U\\1", text$V3, perl=TRUE)

text$V3 <- gsub("AQM", "Aquatic Marine", text$V3, perl=TRUE)
text$V3 <- gsub("AQF", "Aquatic Freshwater", text$V3, perl=TRUE)
text$V3 <- gsub("AQB", "Aquatic Brackish", text$V3, perl=TRUE)

coarse <- c("Aerosol", "Aquatic", "", "", "", "", "", "", "", "", "Built", "", "", "", "", "Plant associated", "", "", "Soil", "", "", "", "Zoological", "", "", "", "Other")

text <- cbind(coarse, text)

colnames(text) <- paste0("\\textbf{", c("Coarse", "Category", "Description"), "}")
composite <- xtable(text)
align(composite) <- "lllp{11cm}"
print(composite, include.rownames=FALSE, include.colnames=TRUE, hline.after = seq(-1, nrow(composite)), comment=FALSE, size="footnotesize")

```