### 2015-12-05 results/tables/build_environmental_categories_table.Rmd
Add table for environmental categories

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
    template: notitle.tex
tables: true
geometry: margin=1in
---

\captionsetup{labelformat=empty}

```{r results='asis', echo=FALSE, warning=FALSE, eval=TRUE}
require(xtable, quietly=TRUE)
options(xtable.sanitize.text.function=identity)

categories <- c(AE="", AQB="Brackish", AQBS="Brackish sediment", AQF="Freshwater", AQFS="Freshwater sediment", AQM="Marine", AQMS="Marine sediment", AQH="Hydrothermal vent", AQI="Ice", AQO="Other", BD="Digesters", BF="Food-associated", BI="Industrial/mining", BP="Pollution associated", BO="Other", PR="Plant root", PS="Plant surface", PO="Other", SA="Agricultural soil", SD="Desert soil", SP="Permafrost", SO="Other", ZV="Vertebrate", ZA="Arthropod", ZN="Other invertebrate", ZO="Other", OT="")

text <- read.table("doc/notebook/ClassificationCode.md", sep=':', row.names=1, stringsAsFactors=FALSE)

text <- text[names(categories),]

text$V2 <- categories

text$V3 <- gsub("^ +", "", text$V3)
text$V3 <- gsub("^(.)", "\\U\\1", text$V3, perl=TRUE)

text$V3 <- gsub("AQM", "Aquatic Marine", text$V3, perl=TRUE)
text$V3 <- gsub("AQF", "Aquatic Freshwater", text$V3, perl=TRUE)
text$V3 <- gsub("AQB", "Aquatic Brackish", text$V3, perl=TRUE)

coarse <- c("\\midrule\\multirow{2}{*}{Aerosol}",
			"\\midrule\\multirow{16}{*}{Aquatic}", "", "", "", "", "", "", "", "",
			"\\midrule\\multirow{11}{*}{Built}", "", "", "", "",
			"\\midrule\\multirow{6}{*}{Plant associated}", "", "",
			"\\midrule\\multirow{6}{*}{Soil}", "", "", "",
			"\\midrule\\multirow{6}{*}{Zoological}", "", "", "",
			"\\midrule\\multirow{2}{*}{Other}")

text <- cbind(coarse, text)

colnames(text) <- paste0("\\textbf{", c("Coarse", "Fine", "Description"), "}")

addtorow <- list(pos=as.list(1:nrow(text)),
				command=c('',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'\\cmidrule(r){2-3}',
						'',
						'')

						)

caption <- "\\textbf{Supplemental Table 1. Description of environmental categories and the criteria used to assign sequences to each category.}"

composite <- xtable(text, caption=caption, label="")
align(composite) <- "lllp{10cm}"
print(composite, include.rownames=FALSE, include.colnames=TRUE, hline.after = c(-1, nrow(composite)), comment=FALSE, add.to.row=addtorow, caption.placement = "top", size="footnotesize")
```