### 2015-12-05 results/tables/build_archaeal_culture_effort_table.Rmd
Add supplemental tables describing culturability

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=1.0in
---

\captionsetup{labelformat=empty}

```{r results='asis', echo=FALSE, warning=FALSE, eval=TRUE}
require(xtable, quietly=TRUE)
options(xtable.sanitize.text.function=identity)

data <- read.table(file="data/process/archaea.cultured_by_time_counts.tsv", header=T, stringsAsFactors=F)

all <- data[,grepl("all_", colnames(data))]
all$all_uncult_nseqs <- all$all_total_nseqs - all$all_cult_nseqs
all$all_uncult_sobs <- all$all_total_sobs - all$all_cult_sobs

all <- all [,c("all_cult_nseqs", "all_uncult_nseqs", "all_total_nseqs",
				"all_cult_sobs", "all_uncult_sobs", "all_total_sobs" )]

all <- all[order(all$all_total_sobs, decreasing=T),]

rownames(all) <- gsub("Candidate_division_", "", rownames(all))
rownames(all) <- gsub("_", " ", rownames(all))

total <- apply(all, 2, sum)
all <- rbind(all, Total=total)

all <- cbind(rownames(all), all)

colnames(all) <- c("", "\\textbf{Cultured}", "\\textbf{Uncultured}", "\\textbf{Total}",  "\\textbf{Cultured}", "\\textbf{Uncultured}", "\\textbf{Total}")
#all <- rbind(header, all)

addtorow <- list(pos=list(-1),
				command=c('\\hline\\multirow{3}{*}{\\textbf{Phylum}} & \\multicolumn{3}{c}{\\textbf{Sequences}} & \\multicolumn{3}{c}{\\textbf{OTUs}} \\\\ \\cmidrule(r){2-4}	\\cmidrule(r){5-7}'))

caption <- "\\textbf{Supplementary Table 7. Frequency that each archaeal sequence or OTU was retrieved by cultivation or by culture-independent methods.}"

composite <- xtable(all, caption=caption, label="")

align(composite) <- paste0('llcccccc')

print(composite, include.rownames=FALSE, include.colnames=TRUE, hline.after = c(0, nrow(composite)-1, nrow(composite)), comment=FALSE, add.to.row=addtorow, caption.placement = "top", size="scriptsize")
```