### 2015-12-07 results/tables/build_archaeal_phylum_effort_table.Rmd

---
output:
  pdf_document:
    includes:
      in_header: table_header.tex
tables: true
geometry: margin=1.0in
---

\captionsetup{labelformat=empty}

```{r results='asis', echo=FALSE, warning=FALSE, eval=TRUE}
require(xtable, quietly=TRUE)
options(xtable.sanitize.text.function=identity)

data <- read.table(file="data/process/archaea.phyla.counts.tsv", header=T)
data <- data[order(data$all, decreasing=T),]
data$ratio_of_ratios <- round(data$ratio_of_ratios, 2)
data$ratio_of_ratios[data$ratio_of_ratios=="Inf"] <- "NA"

rownames(data) <- gsub("\\(.*\\)", "", rownames(data))
rownames(data) <- gsub("_", " ", rownames(data))

total <- apply(data[, 1:3], 2, sum)
data <- rbind(data, Total=c(total, "1.00"))

data <- cbind(rownames(data), data)
colnames(data) <- c("\\textbf{Phylum}",  "\\textbf{<2006 and Before}", "\\textbf{After 2006}", "\\textbf{Total}", "\\textbf{Ratio of Ratios}")

caption <- "\\textbf{Supplementary Table 5. Frequency that each archaeal phylum was sequenced before and after 2006.}"

composite <- xtable(data, caption=caption, label="")

align(composite) <- paste0('ll', paste(rep('c', ncol(data)-1), collapse=""))

print(composite,
	include.rownames=FALSE,
	include.colnames=TRUE,
	hline.after = c(-1,0, nrow(composite)-1,nrow(composite)),
	caption.placement = "top",
	comment=FALSE,
	table.placement="!ht",
	size="normalsize")

```